DATA 621 - Business Analytics and Data Mining

### 2020-09-24 HW2/HW2-Group2.rmd
Finished 1–4 starting 5

---
title: 'DATA 621 - Business Analytics and Data Mining'
subtitle: 'Fall 2020 - Group 2 - Homework #2'
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
    Bryan Persaud, Jeff Shamp
date: "10/11/2020"
output:
  html_document: default
---

@@

The raw confusion matrix 
```{r data_table}
class_data = data %>% select(class, pred_class = scored.class,
                             pred_prob = scored.probability) 
class_data %>% head() %>% knitr::kable(caption = "Preview of the Dataset")
```

### Accuracy of the Predictions
The function, `accuracy` below takes as input a data frame whose first two
columns are the true class and the modeled class respectively. It returns the
accuracy of the predictions using the following formula:

@@

```{r accuracy, echo = TRUE}
accuracy <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt = table(df[, 1:2])
    sum(diag(dt)) / sum(dt)
  }
}
```

```{r accuracy_check}
sprintf("The accuracy of the predictions for the classification data set is %1.3f%%.",
        accuracy(class_data) * 100)
```

### Error Rate of the Predictions
Function `error` that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the classification error rate of the predictions. The formula is:

\[\text{Classification Error Rate} = \frac{FP + FN}{TP + FP + TN + FN}\]

```{r error, echo = TRUE}
error = function(df) {
  if (is.data.frame(df) == FALSE) {
    return("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt = as.data.frame(table(df[, 1:2]))
    FP = dt %>% filter(class == 0 & pred_class == 1) %>% select(Freq)
    FN = dt %>% filter(class == 1 & pred_class == 0) %>% select(Freq)
    return((FP + FN) / nrow(df))
  }
}
```

```{r error_check}
sprintf("The error rate of the predictions for the classification data set is %0.3f.", error(class_data))

sprintf("The accuracy and error rate of the predictions must sum to 1. For the classification data set, there summation equals to %.3f.", accuracy(class_data) + error(class_data))
```

### Precision of the Predictions

Function `precision` that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the precision of the predictions. The formula is:

\[precision = \frac{TP}{TP + FP}\]

>>>>

---
title: "DATA 621 - Business Analytics and Data Mining"
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg, Bryan Persaud,
  Jeff Shamp
date: "10/11/2020"
output:
  html_document: default
  pdf_document: default
subtitle: 'Fall 2020 - Group 2 - Homework #2'
---

@@

```{r data_table}
class_data = data %>% select(class, pred_class = scored.class,
                             pred_prob = scored.probability) 
table(class_data[,1:2]) %>% knitr::kable(caption = "Raw Confusion Matrix")
```

The raw confusion matrix is shown above. In this case, the rows represent the 
**true** classes and the columns represent the **predicted** classes. This is
not the same order as the hand-drawn confusion matrix above, but both are
internally consistent.

# 3. Accuracy
The function `ACC` below takes as input a data frame whose first two columns are
the true class and the modeled class respectively. It returns the accuracy of
the predictions using the following formula:

@@

```{r accuracy, echo = TRUE}
ACC <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt <- table(df[, 1:2])
    sum(diag(dt)) / sum(dt)
  }
}
```

For this dataset the accuracy is 
`r prettyNum(ACC(class_data) * 100, digits = 5L)`%.

# 4. Classification Error Rate
The function `CER` below takes as input a data frame whose first two columns
are the true class and the modeled class respectively. It returns the
classification error rate of the predictions using the following formula:

\[
\textrm{Classification Error Rate} = \frac{FP + FN}{TP + FP + TN + FN}
\]

```{r error, echo = TRUE}
CER <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt <- table(df[, 1:2])
    (sum(dt) - sum(diag(dt))) / sum(dt)
  }
}
```

For this dataset the classification error rate is
`r prettyNum(CER(class_data) * 100, digits = 5L)`%.

Mathematically, the classification error rate must be identical to
\(1 - \textrm{Accuracy}\). In this data set that is clear as 
`r prettyNum(ACC(class_data) * 100, digits = 5L)`% +
`r prettyNum(CER(class_data) * 100, digits = 5L)`% = 
`r prettyNum((CER(class_data) + ACC(class_data)) * 100, digits = 5L)`%.

# 5. Precision
The function `PRC` below takes as input a data frame whose first two columns
are the true class and the modeled class respectively. It returns the
classification error rate of the predictions using the following formula:

\[
\textrm{Precision} = \frac{TP}{TP + FP}
\]
